Supplemental Materials: \emph{Simultaneous segmentation and registration of diffusion MR images of the brain driven by active-contours}

### 2015-03-27 Publications/2015-NeuroImage/03-introduction.tex
update intro

To solve registration, \cite{saad_new_2009} used Pearson correlation to perform linear
  alignment of \gls*{t1} and \emph{b0}.
Another linear registration method is \emph{bbregister} \citep{greve_accurate_2009},
  that uses active contours extracted from the \gls*{t1}\footnote{using \emph{FreeSurfer}
  \citep{fischl_freesurfer_2012}} to look for intensity boundaries in the \emph{b0}
  image.
In practice, the target \emph{b0} does not present clear frontiers of intensity between
  interfacing tissues, what limits \emph{bbregister} to mapping the cortical surface only.
Since the distortions found in \gls*{dmri} are nonlinear, \emph{bbregister} excludes
  from the boundary search those regions typically warped by the artifact.
This tool has become the standard method for its proven robustness.
Nonlinear registration has been successfully performed between \gls*{t2} and \emph{b0}
  images for their similarity, but uniquely as a way to correct for distortions
  \citep{kybic_unwarping_2000,studholme_accurate_2000,wu_comparison_2008,tao_variational_2009}.
Further registration of \gls*{t1} and \gls*{t2} images was still required to map the anatomical
  information (and the \gls*{wm} segmentation) into \gls*{dmri} space.

@@

The hypothesis under this research is that the registration and segmentation
  problems can be solved jointly, increasing the geometrical accuracy of the process.
We propose \emph{regseg} as a joint solution to distortion correction and segmentation,
  using active contours without edges \citep{chan_active_2001} that evolve driving a
  free-deformation field of B-Spline functions.
Optimization is performed using a descent strategy of shape-gradients
  \citep{herbulot_segmentation_2006,besson_dream2s_2003}, therefore \emph{regseg}
  does not implement level sets as most of the presented methods.
Given the properties of the distortion, \emph{regseg} includes the anisotropic regularization
  of displacement fields of \cite{nagel_investigation_1986}.

% While all the correction methods require an extra acquisition, \emph{regseg} is especially
%   indicated in scanning protocols without specific data for correction,
%   such as historical datasets.
Finally, we present an instrumentation framework to evaluate correction and
  segmentation pipelines using realistic distortions as ground truth.
With the integration of \emph{regseg} and an in-house implementation of the 
  \gls*{t2b} correction in the framework, we compare the performance
  of both solutions.

>>>>

To solve registration, \cite{saad_new_2009} used Pearson correlation to perform linear
  alignment of \gls*{t1} and \emph{b0}.
Another linear registration method is \emph{bbregister} \citep{greve_accurate_2009},
  uses active contours with edges to look for intensity boundaries in the \emph{b0}
  image.
The active contours are initialized by surfaces extracted from the 
  \gls*{t1}\footnote{using \emph{FreeSurfer}\citep{fischl_freesurfer_2012}}:
  the pial surface (exterior of cortical \gls*{gm}) and the \gls*{wm}/\gls*{gm} interface.
In practice, the \emph{b0} image only presents a detectable frontier for the pial surface
  and \emph{bbregister} is limited to align the cortical surface.
Since the distortions found in \gls*{dmri} are nonlinear, \emph{bbregister} excludes
  from the boundary search those regions typically warped by the artifact.
This tool has become the standard method for its proven robustness.
Nonlinear registration has been successfully performed between \gls*{t2} and \emph{b0}
  images for their similarity, but uniquely as a way to correct for distortions
  \citep{kybic_unwarping_2000,studholme_accurate_2000,wu_comparison_2008,tao_variational_2009}.
Further registration of \gls*{t1} and \gls*{t2} images was still required to map the anatomical
  information (and the \gls*{wm} segmentation) into \gls*{dmri} space.

@@

The hypothesis under this research is that the registration and segmentation
  problems in \gls*{dmri} can be solved simultaneously, increasing the geometrical
  accuracy of the process.
We propose a tool called \emph{regseg} to exploit the prior information of shapes
  extracted on \gls*{t1} images to register the anatomical reference
  to \gls*{dmri} space, implicitly segmenting these data.
The approach uses active contours without edges \citep{chan_active_2001} that evolve driving a
  free-deformation field of B-Spline functions.
Optimization is performed using a descent strategy of shape-gradients
  \citep{besson_dream2s_2003,herbulot_segmentation_2006}, therefore \emph{regseg}
  does not implement levelsets as most of the presented methods.
The nonlinear warping caused by distortion is aligned with one of the imaging axis due
  to the theory underpinning \gls*{epi} sequences.
Therefore, \emph{regseg} includes an anisotropic regularizer of the displacement field 
  proposed by \cite{nagel_investigation_1986}.
Finally, we present an instrumentation framework to evaluate correction and
  segmentation pipelines using realistic distortions as ground truth.
With the integration of \emph{regseg} and an in-house implementation of the 
  \gls*{t2b} correction in the framework, we compare the performance
  of both solutions.
